### 2017-06-19 manuscript/manuscript.Rmd
Everything is code; contrast VCS to "standard workflow"

Given this broad definition of reproducibility, it is easy to recognize that many projects fall between the extremes of completely reproducible and not reproducible at all. A completely reproducible project would provide consumers the entire raw data set, and detailed, unambiguous instructions to analyze it (usually as computer code). This degree of reproducibility is very rare in Psychology [@wicherts_poor_2006; @vanpaemel_are_2015], although data sharing has become more common in recent years in journals that award authors who do so with special badges [@kidwell_badges_2016]. However, even when authors (report to) share data, the shared data sets sometimes turn out to be incorrect, not usable, incomplete, or not available at all [@kidwell_badges_2016]. This suggests that reproducibility requires more than good intentions. What, then, are some specific challenges to reproducibility?

@@

Version Control Systems are computer programs designed for tracking changes to computer code, and collaborating on the code with others. Although initially developed for writing code collaboratively[^vcs-collaborate], it is easy to recognize that VCS can also be adopted for scientific production and collaboration. For example, behavioral experiments are often created with a programming language, and can have multiple authors and versions. Keeping track of the versions of the program and changes to the code, and allowing many authors to contribute to it (without breaking the experimental program) are problems that VCSs were specifically designed for. VCS is also easily adopted to other aspects of the research cycle, such as curating data across computers and laboratories, or writing manuscripts with multiple authors, versions, and sources of results.

@@

Some popular Version Control Systems are SVN, Mercurial, and Git. In this tutorial, we focus on Git: Git is already incresing in popularity in the scientific community, and is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.

@@

@@

[^git-version]: As of the writing of this article, the current version of Git is 2.13.1.
[^git-name]: The creator of Git, Linus Torvalds, named Git after himself as "the stupid content tracker" (@mcmillan_after_2005; <https://git-scm.com/docs/git.html>)
[^simple]: By "simple", we only mean that the technical aspects of the project, related to curating materials related to it, are simple. We of course do not suggest that there might be anything theoretically or scientifically simple in such projects.
[^vcs-collaborate]: The software we present below is used used by major software developers such as Microsoft, Google, and Facebook on code bases with hundreds of contributors.
[^cli-links]: If using a text-based command line seems challenging, Codecademy (<https://www.codecademy.com/learn/learn-the-command-line>) has a free interactive online tutorial, and MIT offers a free online game to teach using the command line (<http://web.mit.edu/mprat/Public/web/Terminus/Web/main.html>).
[^git-link-submodules]: For advanced users, Git *submodules* allow linking projects to each other, or organizing more complex projects into projects and their sub-projects (<https://git-scm.com/book/en/v2/Git-Tools-Submodules>).
[^git-cheat-sheet]: It helps to have a Git command cheat sheet (<https://services.github.com/resources/cheatsheets/>) printed and taped on your wall, but it contains many more commands than are needed for the basic use of Git in standard Psychology studies.
[^tier-link]: <http://www.projecttier.org/tier-protocol/specifications/>
[^git-branches]: A particularly powerful approach for trying new features is **branching**: The project can be duplicated to a new branch and modified, then merged back to the main branch after work on the new feature is complete---or the new branch can be discarded if the work ended unsatisfactory. Branches are outside the scope of this tutorial, for more information see the Git website (<https://git-scm.com/book/en/v2/Git-Branching-Branches-in-a-Nutshell>).

>>>>

Given this broad definition of reproducibility, it is easy to recognize that many projects fall between the extremes of completely reproducible and not reproducible at all. A completely reproducible project would provide consumers the entire raw data set with detailed and unambiguous instructions to analyze it (usually as computer code). This degree of reproducibility is very rare in Psychology [@wicherts_poor_2006; @vanpaemel_are_2015], although data sharing has become more common in recent years in journals that award authors who do so with special badges [@kidwell_badges_2016]. However, even when authors (report to) share data, the shared data sets sometimes turn out to be incorrect, not usable, incomplete, or not available at all [@kidwell_badges_2016]. This suggests that reproducibility requires more than good intentions. What, then, are some specific challenges to reproducibility?

@@

Version Control Systems (VCS) are computer programs designed for tracking changes to computer code, and collaborating on the code with others. VCS were initially developed by software engineers for writing code collaboratively[^vcs-collaborate], but are increasingly being adopted to enhance workflows outside computer science. To help understand why, it is helpful to think of "code" more broadly as any text written on a computer: Manuscripts, books, statistical analysis scripts, source code of computerized experiments, and even data files are "code" or have source code, which is just plain text written on a computer, and which can benefit from version control. Going forward, when we use the word "code" in this tutorial, we mean it in this broad sense (e.g. this manuscript's source "code" was written on a computer, and was version controlled.)

For example, computerized behavioral experiments are written with a computer programming language such as MATLAB or Python. The experiment's source code (text written by human, interpreted by computer to e.g. display stimuli to participants) may have multiple authors and usually goes through multiple versions. Keeping track of the versions of the program and changes to the code, and allowing many authors to contribute to it (without breaking the experimental program) are problems that VCSs were specifically designed for. It might be less obvious that writing a manuscript is quite similar, as far as the computer is concerned: Multiple authors write multiple versions of a text document, and sometimes previous versions need to be inspected, and the text needs to be "merged" across the many authors. Even less obvious is that data itself is often plain text: In most computerized behavioral experiments, the output data are numbers and text written into a text file. These text files can then be version controlled (and researchers would usually not want to see that their raw data files have changed after they were created---VCS allows verifying that they haven't changed because their history is logged.)

@@

To understand what VCS is, it is useful to contrast an example common case with VCS and what we believe is a fairly standard workflow. Consider collaborating with one person on a manuscript that reports results from a data set using some statistical model. In the standard workflow, one person might organize the raw data in one way to fit a statistical model, and then write a draft of the manuscript. This would create three files: `data.csv`[^csv], `analysis.R`, and `manuscript.doc`. If the coauthor then wished to explore another statistical model, which requires the data in a different format, and then edit the manuscript, she would create three more files: `data_new.csv`[^csv], `analysis_new.R`, and `manuscript_new.doc`. This cycle would then repeat as many times as required, each time creating more files, making it more and more difficult for the authors to exactly remember which data was paired with which analysis, and which analyses (and data format) was reported in which version of the paper.

Using VCS the workflow is greatly simplified: There are only three files (data, analysis, and manuscript), but they are all under version control, which keeps track of changes to them. Because changes are monitored, creating new files for every new idea or edit is unnecessary, thereby possibly reducing room for error in remembering which data file was linked with which analysis, and which manuscript version had the correct numerical results, and so forth. Further, as we explain below, VCS allows many coauthors to work on the files simultaneously, eschewing the need for emailing a version to a coauthor, then waiting for them to make changes before continuing with your work. Box 1 has further practical examples of where VCS can be used to improve the scientific workflow.

@@

Version control software has a long history in software engineering, and there are many VCS programs. Some popular ones are Apache Subversion[^subversion], Mercurial[^mercurial], and Git. In this tutorial, we focus on Git: Git is already incresing in popularity in the scientific community, and is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.

@@

[^csv]: `.csv` stands for comma separated values, and is a plain text file commonly used to store two-dimensional data.
[^git-version]: As of the writing of this article, the current version of Git is 2.13.1.
[^git-name]: The creator of Git, Linus Torvalds, named Git after himself as "the stupid content tracker" (@mcmillan_after_2005; <https://git-scm.com/docs/git.html>)
[^simple]: By "simple", we only mean that the technical aspects of the project, related to curating materials related to it, are simple. We of course do not suggest that there might be anything theoretically or scientifically simple in such projects.
[^vcs-collaborate]: The software we present below is used used by major software developers such as Microsoft, Google, and Facebook on code bases with hundreds of contributors.
[^cli-links]: If using a text-based command line seems challenging, Codecademy (<https://www.codecademy.com/learn/learn-the-command-line>) has a free interactive online tutorial, and MIT offers a free online game to teach using the command line (<http://web.mit.edu/mprat/Public/web/Terminus/Web/main.html>).
[^git-link-submodules]: For advanced users, Git *submodules* allow linking projects to each other, or organizing more complex projects into projects and their sub-projects (<https://git-scm.com/book/en/v2/Git-Tools-Submodules>).
[^git-cheat-sheet]: It helps to have a Git command cheat sheet (<https://services.github.com/resources/cheatsheets/>) printed and taped on your wall, but it contains many more commands than are needed for the basic use of Git in standard Psychology studies.
[^tier-link]: <http://www.projecttier.org/tier-protocol/specifications/>
[^git-branches]: A particularly powerful approach for trying new features is **branching**: The project can be duplicated to a new branch and modified, then merged back to the main branch after work on the new feature is complete---or the new branch can be discarded if the work ended unsatisfactory. Branches are outside the scope of this tutorial, for more information see the Git website (<https://git-scm.com/book/en/v2/Git-Branching-Branches-in-a-Nutshell>).
[^subversion]: <https://subversion.apache.org/>
[^mercurial]: <https://www.mercurial-scm.org/>
